Boosting Higgs Pair Production in the $b\bar{b}b\bar{b}$ Final State with Multivariate Techniques

### 2015-12-08 paper/sec-intro.tex
More work on paper

In the context of SM production,
the main advantage of the $b\bar{b}b\bar{b}$ final state is the
enhancement of the signal yield
from the large branching fraction of Higgs bosons into bottom-anti-bottom
pairs, ${\rm BR}\lp H\to b\bar{b}\rp\simeq 0.57$~\cite{Dittmaier:2012vm}.
%
However, a measurement in this channel
needs to deal with an overwhelming QCD multi-jet background.
%
Recent studies of Higgs pair production in this
final state~\cite{Wardrope:2014kya,deLima:2014dta}
estimate that, for an integrated
luminosity of
$\mathcal{L}$=3000 fb$^{-1}$,
a signal significance of around $S/\sqrt{B}\sim 2.0$ could be obtained.
%
While irreducible backgrounds, such as $4b$ and
$t\bar{t}$, are included in these analysis, the contribution
from reducible processes, such as $bbjj$ and
$jjjj$, which can contribute to the signal yield when 
light jets are mis-identified as $b$-jets,
are not accounted for.
%
As will show, due to both
selection effects and $b$-quark radiation in the
parton shower, the
contribution of the $2b2j$ process is actually comparable to
the irreducible $4b$ component.

In this work, we revisit the feasibility of SM Higgs pair production in
gluon-fusion
in the $b\bar{b}b\bar{b}$ final state at the LHC.
%
 Our strategy is based on a combination of traditional cut-based
 methods with multivariate analysis (MVA).
  %
  We account for  all relevant
  backgrounds, including the contribution from mis-identified
  light and charm jets.
  %
  We also assess the robustness of our analysis strategy in
  an environment with high pile-up (PU).
  %
  We find an overall signal significance of $S/\sqrt{B}\simeq 4.0$
  for a total integrated
  luminosity of $\mathcal{L}$=3000 fb$^{-1}$,
   indicating
  that the $b\bar{b}b\bar{b}$ channel
alone should allow for the observation of double Higgs production
  at the HL-LHC.
%

>>>>

In the context of SM production,
the main advantage of the $b\bar{b}b\bar{b}$ final state is the
enhancement of the signal yield
from the large branching fraction of Higgs bosons into bottom-anti-bottom
pairs, ${\rm BR}\lp H\to b\bar{b}\rp\simeq 0.57$~\cite{Dittmaier:2012vm}.
%
However, a measurement in this channel
needs to deal with an overwhelming QCD multi-jet background.
%
Recent studies of Higgs pair production in this
final state~\cite{Wardrope:2014kya,deLima:2014dta}
estimate that, for an integrated
luminosity of
$\mathcal{L}=3$ ab$^{-1}$,
a signal significance of around $S/\sqrt{B}\sim 2.0$ could be obtained.
%
While irreducible backgrounds, such as $4b$ and
$t\bar{t}$, are included in these analysis, the contribution
from reducible processes, such as $bbjj$ and
$jjjj$, which can contribute to the signal yield when 
light jets are mis-identified as $b$-jets,
are not accounted for.
%
As will show, due to both
selection effects and $b$-quark radiation in the
parton shower, the
contribution of the $2b2j$ process is actually comparable to
the irreducible $4b$ component.

In this work, we revisit the feasibility of SM Higgs pair production in
gluon-fusion
in the $b\bar{b}b\bar{b}$ final state at the LHC.
%
 Our strategy is based on a combination of traditional cut-based
 methods with multivariate analysis (MVA).
  %
  We account for  all relevant
  backgrounds, including the contribution from mis-identified
  light and charm jets.
  %
  We also assess the robustness of our analysis strategy in
  an environment with high pile-up (PU).
  %
  We find an overall signal significance of $S/\sqrt{B}\simeq 4.0$
  for a total integrated
  luminosity of $\mathcal{L}=3$ ab$^{-1}$,
   indicating
  that the $b\bar{b}b\bar{b}$ channel
alone should allow for the observation of double Higgs production
  at the HL-LHC.
%

### 2015-12-08 paper/sec-results.tex
More work on paper

In this section we present the results of the 
cut-based analysis, including the corresponding
cut-flow.
We
compare our results with other recent studies,
and in particular
we study how the signal significance
is affected if only the $4b$ component of the
QCD multi-jet background is taken into account,
while the $2b2j$ and $4j$ components are neglected.
%

@@

First of all, we compare the cross-sections at various
stages of the selection
for the three categories, for signal and background events,
by providing a detailed cut-flow.
%
We consider all backgrounds enumerated in Sect.~\ref{mcgeneration},
and also discuss how results are modified if only the $4b$
component is kept.
%
At each stage of the cut-flow, we also provide the
the signal significance, $S/\sqrt{B}$, and the signal
over background ratio, $S/B$, corresponding to the
 HL-LHC with
an integrated luminosity of $\mathcal{L}
=3000$ fb$^{-1}$.
%
For this type of analysis,
it is crucial not only to achieve a good signal
significance, $S/\sqrt{B}$, but also a good signal over background ratio $S/B$.
%
If the latter is too small, an unrealistically precise
measurement of the background would be required.
%
In Table~\ref{tab:cutflowdetails}
we describe this cut-flow,
separated into the (exclusive) boosted, intermediate,
    and resolved categories.
    %
    From top to bottom, additional cuts are required in addition
    to the previous ones.
    %
    The various cuts
    are defined as follows:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \begin{tabular}{|c|c|c|c|}
\hline
&  Boosted  &   Intermediate &  Resolved  \\
\hline
\hline
{\bf C0} &  \multicolumn{3}{c|}{Generator level} \\
\hline
{\bf C1a} & $N_{\rm jets}^{R10}\ge 2$ & $N_{\rm jets}^{R04}\ge 2$, $N_{\rm jets}^{R10}=1$  &
$N_{\rm jets}^{R04}\ge 4$ \\
\hline
{\bf  C1b} & \multicolumn{3}{c|}{+$p_T$ cuts} \\
{\bf C1c} & \multicolumn{3}{c|}{+rapidity cuts}\\
\hline
 {\bf C1d} & +$N_{\rm MDT}\ge 2$ & +$N_{\rm jets}^{R10}=1$ with MDT  &
 +Higgs reconstruction \\
 \hline
{\bf C1e} & \multicolumn{3}{c|}{ +$m_H$ window cut} \\
\hline
{\bf C2} & \multicolumn{3}{c|}{+$b$-tagging}    \\
\hline
  \end{tabular}
  \caption{\small Cut-flow of the present analysis for the boosted, intermediate
    and resolved selections.
    %
      \label{tab:cutflowdetails}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{itemize}
    \item {\bf C0}: these are the generator level cross-sections, with
      mild generation cuts in the case of the background processes, and no
      generation cuts
      for the signal events.
      %
    \item {\bf C1a}:  we only check that we have at least
      two large-$R$ jets (in the boosted case),
      one large-$R$ jet and at least 2 small-$R$ jets (in the intermediate
      case) and at least four small-$R$ jets (in the resolved case).
      %
      No acceptance requirements are imposed.
    \item {\bf C1b}: require that the above jets should
      satisfy the corresponding $p_T$ thresholds,
      $p_T \ge 200$ GeV for large-$R$ jets and
      $p_T \ge 50$ GeV for small-$R$ jets.
    \item {\bf C1c}: same for the
      rapidity acceptance requirement.
    \item {\bf C1d}: the two leading large-$R$ jets must
      be mass-drop tagged in the boosted category.
      %
      In the intermediate category, the large-$R$ jet must also be mass-drop tagged.
      %
    \item {\bf C1e}: after the two Higgs candidates  have been reconstructed,
      their invariant masses are required to lie a window around $m_H$,
      in particular between 85 and 165 GeV.
          \item {\bf C2}: finally, the
            $b$-tagging conditions are
            imposed, see
            Sect.~\ref{sec:btagging}.
      \end{itemize}
    Signal and background events satisfying all the analysis cut up to the
    C2 level
    are then used as input for the MVA training, to be described next
    in Sect.~\ref{sec:mva}.

    %
    In Table~\ref{tab:cutflow_noPU_1} we provide
    the values of cross-sections, in femtobarns,
     for the signal
      and the various background
      components as a function of the
      cut-flow, for the resolved,
      boosted and intermediate
      categories.
      %
       In each case, we also provide the signal over
      background ratio, $S/B$, and the signal
      significance, $S/\sqrt{B}$, at the HL-LHC, considering either
      the total background or only the $4b$ component.
     %
      This separation emphasizes
      the  importance of the reducible component of the QCD background.
      %
      Even after $b$-tagging, the reducible $2b2j$ background
      is comparable, or even larger, than the irreducible
      $4b$ component.
      %
      %
      So at the end of the cut-based analysis, signal significance
      are overestimated by up to a factor 2 if the light
      and charm jet background is neglected.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \scriptsize
  \input{tables/table_noPU_1.tex}
  $\,$ \\
  \vspace{0.5cm}
  \input{tables/table_noPU_2.tex}
  $\,$ \\
  \vspace{0.5cm}
    \input{tables/table_noPU_3.tex}
    \caption{\small The cross-sections, in femtobarns,
      for the signal and the various background
      processes at different steps of the
      cut-flow, for the resolved (upper table),
      intermediate (middle table) and boosted
      (lower table) categories, for the analysis
      without PU.
      %
      In each case, we also provide the signal over
      background ratio, $S/B$, and the signal
      significance, $S/\sqrt{B}$, considering either
      the total background or only the $4b$ component.
      %
      The different levels of the cut-flow are summarized
      in Table~\ref{tab:cutflowdetails}.
 \label{tab:cutflow_noPU_1}}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

In this section we present the results of the 
cut-based analysis and provide a
cut-flow for the different analysis steps.
%
We
compare our results with other recent studies,
and assess how the signal significance
is affected if only the $4b$ component of the
QCD multi-jet background is taken into account but not 
the $2b2j$ and $4j$ components.
%
We present now the results in the no PU scenario, the impact
of PU will be discussed in the next section.

@@

We now compare the cross-sections for
signal and background events at various
stages of the analysis.
%
We consider all relevant backgrounds (see Sect.~\ref{mcgeneration}),
and discuss how results are modified if only the $4b$
component is kept.
%
At each stage of the cut-flow, we also provide the
the signal significance, $S/\sqrt{B}$, and the signal
over background ratio, $S/B$, corresponding to the
 HL-LHC with
an integrated luminosity of $\mathcal{L}
=3000$ fb$^{-1}$.

In Table~\ref{tab:cutflowdetails}
we summarize the definition of the different
levels of the cut-flow of the present analysis,
separated into the boosted, intermediate,
    and resolved categories.
    %
    From top to bottom, additional cuts are required in addition
    to the previous ones.
    %
    In more details, these different analysis levels are the following:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \begin{tabular}{|c|c|c|c|}
\hline
&  Boosted  &   Intermediate &  Resolved  \\
\hline
\hline
{\bf C0} &  \multicolumn{3}{c|}{Generator level} \\
\hline
{\bf C1a} & $N_{\rm jets}^{R10}\ge 2$ & $N_{\rm jets}^{R04}\ge 2$, $N_{\rm jets}^{R10}=1$  &
$N_{\rm jets}^{R04}\ge 4$ \\
\hline
{\bf  C1b} & \multicolumn{3}{c|}{+$p_T$ cuts} \\
{\bf C1c} & \multicolumn{3}{c|}{+rapidity cuts}\\
\hline
 {\bf C1d} & +$N_{\rm MDT}\ge 2$ & +$N_{\rm jets}^{R10}=1$ with MDT  &
 +Higgs reconstruction \\
 \hline
{\bf C1e} & \multicolumn{3}{c|}{ +$m_H$ window cut} \\
\hline
{\bf C2} & \multicolumn{3}{c|}{+$b$-tagging}    \\
\hline
  \end{tabular}
  \caption{\small Definition of the cut-flow levels for the boosted, intermediate
    and resolved selections.
    %
      \label{tab:cutflowdetails}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{itemize}
    \item {\bf C0}:  generator level cross-sections, with
      mild generation cuts in the case of the background processes, and no
      generation cuts
      for the signal events.
      %
    \item {\bf C1a}:  check that we have at least
      two large-$R$ jets (in the boosted case),
      one large-$R$ jet and at least 2 small-$R$ jets (in the intermediate
      case) and at least four small-$R$ jets (in the resolved case).
      %
      No acceptance requirements are imposed.
    \item {\bf C1b}: require that the above jets should
      satisfy the corresponding $p_T$ thresholds,
      $p_T \ge 200$ GeV for large-$R$ jets and
      $p_T \ge 50$ GeV for small-$R$ jets.
    \item {\bf C1c}: same for the
      rapidity acceptance requirement.
    \item {\bf C1d}: the two leading large-$R$ jets must
      be mass-drop tagged in the boosted category.
      %
      In the intermediate category, the large-$R$ jet must also be mass-drop tagged.
      %
    \item {\bf C1e}: after the two Higgs candidates  have been reconstructed,
      their invariant masses are required to lie a window around $m_H$,
      in particular between 85 and 165 GeV.
          \item {\bf C2}: the
            $b$-tagging conditions are
            imposed, see
            Sect.~\ref{sec:btagging}.
      \end{itemize}
    Signal and background events satisfying all the analysis cut up to the
    C2 level
    are then used as input for the MVA training, to be described next
    in Sect.~\ref{sec:mva}.
    %
    In Table~\ref{tab:cutflow_noPU_1} we collect
    the values for cross-sections for signal and background processes
    for different levels of the cut flow.
    %
    Results are divided into the resolved, intermediate and boosted categories,
    and are inclusive up to the C2 level, where exclusivity is imposed.
    %

In Table~\ref{tab:cutflow_noPU_1} we also  provide the signal over
      background ratio, $S/B$, and the signal
      significance, $S/\sqrt{B}$, corresponding to an integrated
      luminosity of $\mathcal{L}=3$ ab$^{-1}$.
      %
      This is done both taking into account all the background components and only
      the $4b$ QCD background.
        %
      We find that after $b$-tagging, the  $2b2j$ component is
      of the same order of magnitude as the $4b$ component.
      %
      This means that the signal significance at the end of the cut-based
      analysis is degraded by about a factor two due to the contribution
      of light and charm jets being mis-identified as $b$-jets.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \scriptsize
  \input{tables/table_noPU_1.tex}
  $\,$ \\
  \vspace{0.5cm}
  \input{tables/table_noPU_2.tex}
  $\,$ \\
  \vspace{0.5cm}
    \input{tables/table_noPU_3.tex}
    \caption{\small The cross-sections, in femtobarns,
      for the signal and the various background
      processes at different steps of the
      cut-flow, for the resolved (upper table),
      intermediate (middle table) and boosted
      (lower table) categories, for the analysis
      without PU.
      %
      In each case, we also provide the signal over
      background ratio, $S/B$, and the signal
      significance, $S/\sqrt{B}$, considering either
      the total background or only the $4b$ component, for
      a total integrated luminosity of $\mathcal{L}=3$ ab$^{-1}$.
      %
      The different levels of the cut-flow are summarized
      in Table~\ref{tab:cutflowdetails}.
 \label{tab:cutflow_noPU_1}}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
